A Flux Scale for Southern Hemisphere 21cm EoR Experiments

### 2013-05-06 ms.tex
Rounding out text edits in Discussion and Conclusions

%CALIBRATION
\subsection{Per-Antenna Gain and Phase Calibration}

@@

% XXX time-dependent calibration is critical.  modify paragraph below

Measurements of ambient temperature versus time near the balun amplifier of a fiducial antenna and near
receiver amplifiers were used to divide out the predicted gain variation versus temperature that these amplifiers
are known to exhibit.  A predicted gain profile was derived using coefficients characterized in
laboratory measurements \citep{parashare_bradley2009} and
confirmed in field measurements \citep{pober_et_al2011} of $H_{\rm balun}=-0.024$ dB/K, $H_{\rm recvr}=-0.045$
dB/K, and $H_{\rm cable}=-0.018$ dB/K for 150m-long cables.  Because cable temperatures were not measured
during these observations, the cable temperature was assumed to be the same as the measured balun temperature.

Antenna delays and amplitudes were found by fitting a point source visibility
model to Centaurus A, Pictor A, and Fornax A.  Each source is imperfectly
modeled by a single point-source, but the solution differences are minimized by
averaging over the three independent solutions. These same calibration
solutions have been successfully applied in \citet{Pober:2013XXX} and \citet{Stefan:2013:XXX}.
for power-spectrum analysis of foregrounds and imaging of Centaurus A, respectively.

@@

% XXX exclusion of minimum uv spacing?

@@

The resulting set of spectra were then calibrated to a model of J2331-416 to remove 
the residual bandpass due to the net effect of the primary beam as discussed in \S \ref{sec:approach}. J2331-416 
\footnote{Here we set the spectrum to $S_{150}=33$Jy, $\alpha=-0.76$} fit to measurements beolow 1GHz \citep{slee1995,kuhr_et_al1981,large_et_al1981,burgess_hunstead2006}.}
was selected for its relatively high brightness, spectral smoothness and large quantity
 of available catalog data. Each source track has a slightly different sample profile
 resulting from different amounts of flagged data at each time and channel. To account
 for these differences in the bandpass calibration we build a set of calibration tracks 
 that match the tracks for each source. Calibration of each source then proceeds with
 an optimally matched calibration spectrum.

@@

\[
\logL = \sum_\nu{\frac{\left(S_\nu - S_{150}\left(\frac{\nu}{150}\right)^\alpha\right)^2}{\Delta S_\nu} \label{eq:catlogL}
\]

@@

\begin{figure}[htbp]
\includegraphics[width=0.49\textwidth]{plots/pic_trace_hist.png}
\includegraphics[width=0.49\textwidth]{plots/1556-466_trace_hist.png}
\caption{
A more detailed view of the posterior probability distribution. On the left
Pictor A, on the right 1556-466. Confidence contours range from 0.2 to 0.8.
Compare with the contours for this source shown in Fig. \ref{fig:SI_contour_3}.
As above, grey for  the fit to catalog data, black for joint catalog and PAPER
fit.
}
\end{figure}
% XXX paragraph too remedial. digest
%In general an experiment measures data $D$ and given some model of the
%statistics of the measurement one usually assigns an error bar defining the
%probability of getting that data, given a model $M$.  Given all the new data,
%and prior data, there will be a new optimal set of model parameters $M$ with
%associated error bars.  In general, these error bars are a parameterization of
%a probability distribution function.  The pdf associated with the data often
%called the likelihood ($\mathcal{L}$), while the one corresponding the model
%error bars is the {\em posterior probability distribution}, $\Prob(M|D)$. They
%are related
%
%\[
%\Prob(M|D) = \frac{\mathcal{L}(D|M) \Prob(M)}{\Prob(D)}
%\] 
%by the factor $\frac{\Prob(M)}{\Prob(D)}$, the prior probability of the model
%$M$ (given past experiments) over the normalization factor, integrating over
%all models.  In general the posterior pdf depends on both the known relation
%between the model parameters and as well as correlations hidden within the
%data.  
%
%% XXX too remedial.  digest to 1 senetence
%Direct calculation of the posterior is difficult, particularly the
%normalization factor, however there is a simple procedure for drawing samples
%from the posterior. In this paper we use a Markov Chain Monte Carlo sampler.
%Given any current location in parameter space, this algorithm draws a new set
%of parameters and computes the relative likelihood increase. If the probability
%difference is larger than a number drawn randomly from the set $(0:1]$, the
%jump is accepted ---a Metropolis-Hastings step. In the limit of an infinity of
%such steps, this process has been shown to converge on the true posterior
%distribution.  In practice it is run for a large number of steps until things
%appear to stabilize.

@@

Inspecting the confidence contours we find that, for the vast majority ($\sim
75$\%), the PAPER measurements confirm the power law extrapolation from higher
frequencies. The remaining 25\% either A) have large PAPER error bars and
therefore provide no new information or B) do not agree with the spectral index
model.  To understand where most measurements fall on this spectrum we
numerically quantify the overall model improvement derived from the addition of
the PAPER data as 
\begin{align}
\textrm{Improvement} &= \left(\textrm{Precision increase}\right) \left(\textrm{Catalog agreement}\right)\\
&= \left(\frac{1}{Area(\textrm{PAPER})} - \frac{1}{Area(\textrm{Cat})}\right) \left( Area(\PAPER \cap \textrm{Cat}) \right)
\end{align}
% XXX This equation is yucky

@@

However, the majority of source 78\% have positive improvement index,
indicating strong confirmation of the extrapolated spectrum (see Figures
\ref{fig:SI_contour_1} - \ref{fig:SI_contour_4}).  Pictor A is in the middle of
this group  with an improvement of 0.39 ---only 14 sources show stronger
confirmation.  The flux model is $S150=$388 $\pm$ 9.4Jy, $\alpha=-0.77\pm0.02$.
% XXX need more detail here.  This is the crux of what is done.
The combination of several independent PAPER data points near 150MHz with
previous catalog measurements results in this estimated uncertainty of 2.4\%.
This uncertainty is dominated by agreement between the 2 Wills measurements
near 600MHz which are accurate to 2\% and the five PAPER measurements, which
are accurate to ~5\%.

% XXX comparison with previous in-band results (danny 2011, williams 2012)

@@

% XXX adapt this paragraph to be a summary of the broad goals attained in this paper
The primary purpose of this observation is to produce a high quality Pictor A
% XXX need to coordinate "primary purpose" here with abstract and intro motivation
spectrum. As a verification measure we have observed many other sources at the
same declination for comparison against previous observations.  Our comparison
metric will be the degree to which the PAPER data are consistent with previous
observations,given a model and, if consistent, they improve on the precision of
our spectral model parameters.

Direct measurements of the Pictor A spectrum are key to correctly setting the
flux scale of PAPER and other EoR experiments. Existing EoR band measurements
are accurate to 20\% implying a 40\% uncertainty in the absolute power spectrum
level.  Here we have provided new measurements of Pictor A that are accurate to
2.4\%;  388 $\pm$ 9.4Jy on the \citet{Baars:1977p9678} scale. The PAPER
measurement has a precision of $<$1\%; the majority of the uncertainty comes in
setting the absolute flux scale from prior data.

We also take advantage of multiple PAPER observations to measure the next  61
brightest, previously cataloged, sources. These sources are distributed over
the full 24 hours of Right Ascension coverage but are within 5 degrees
declination of Pictor A.  Using a Bayesian analysis we conclude that most of
these are consistent with previous measurements and provide useful new
constraints. However, in the remainder, the simple spectral model fails to
adequately describe the data. In most cases the proximity to bright or extended
structures implies side lobe contamination.  Therefore, though we include the
model parameters, they are not meant to be definitive, but provide confidence
that the reported PAPER flux values are on the correct global scale.

The data  include new measurements of Pictor A that are uncertain to 3.5\%,
close to the level of the most precise higher frequency measurements by
\citet{Wills:1975p9314}. Using these new measurements we update the spectral
calibration model to an accuracy of 2.5\%. As validation of this method, we
also measure a selection of 58 additional sources at the same Declination.

% XXX looking to the future, what next

>>>>

%CALIBRATION
\subsection{Per-Antenna Gain Calibration}

@@

Small gain variations between antennae and over time introduce significant systematic effects which degrade 
instrument performance. Temporal variations are dominated by changes in ambient temperature. 
Measurements of ambient temperature versus time near the balun amplifier of a fiducial antenna and near
receiver amplifiers were used to divide out the predicted gain variation versus temperature that these amplifiers
are known to exhibit.  A predicted gain profile was derived using coefficients characterized in
laboratory measurements \citep{parashare_bradley2009} and
confirmed in field measurements \citep{pober_et_al2011} of $H_{\rm balun}=-0.024$ dB/K, $H_{\rm recvr}=-0.045$
dB/K, and $H_{\rm cable}=-0.018$ dB/K for 150m-long cables.  Because cable temperatures were not measured
during these observations, the cable temperature was assumed to be the same as the measured balun temperature.

Matching of relative gains and phases between antenna was done by fitting a per-antenna complex gain to portions
of data which have easily modeled sky. 
This calibration is only relative between antennae, absolute flux calibration
comes after time and frequency averaging in  \S \ref{sec:flux_scale}.   
The antenna delays and amplitudes were found by fitting a point source visibility
model to Centaurus A, Pictor A, and Fornax A.  Each source is imperfectly
modeled by a single point-source, but the solution differences are minimized by
averaging over the three independent solutions. These same calibration
solutions have been successfully applied in \citet{Pober:2013XXX} and \citet{Stefan:2013:XXX}.
for power-spectrum analysis of foregrounds and imaging of Centaurus A, respectively.

@@

@@

The resulting set of spectra were then calibrated to a model of J2331-416 to remove 
the residual bandpass due to the net effect of the primary beam as discussed in \S \ref{sec:approach}. J2331-416 
\footnote{Here we set the spectrum to $S_{150}=33$Jy, $\alpha=-0.76$} fit to measurements beolow 1GHz \citep{slee1995,kuhr_et_al1981,large_et_al1981,burgess_hunstead2006}.
was selected for its relatively high brightness, spectral smoothness and large quantity
 of available catalog data. Each source track has a slightly different sample profile
 resulting from different amounts of flagged data at each time and channel. To account
 for these differences in the bandpass calibration we build a set of calibration tracks 
 that match the tracks for each source. Calibration of each source then proceeds with
 an optimally matched calibration spectrum.

@@

\begin{equation}
\logL = \sum_\nu{\frac{\left(S_\nu - S_{150}\left(\frac{\nu}{150}\right)^\alpha\right)^2}{\Delta S_\nu} }
%\label{eq:catlogL}
\end{equation}

@@

\begin{figure}[htbp]
\includegraphics[width=0.49\textwidth]{plots/pic_trace_hist.png}
\includegraphics[width=0.49\textwidth]{plots/1556-466_trace_hist.png}
\caption{
A more detailed view of the posterior probability distribution. On the left
Pictor A, on the right 1556-466. Confidence contours range from 0.2 to 0.8.
Compare with the contours for this source shown in Fig. \ref{fig:SI_contour_3}.
As above, grey for  the fit to catalog data, black for joint catalog and PAPER
fit.
}
\end{figure}

@@

Inspecting the confidence contours we find that, for the vast majority ($\sim
75$\%), the PAPER measurements confirm the power law extrapolation from higher
frequencies. The remaining 25\% either A) have large PAPER error bars and
therefore provide no new information or B) do not agree with the spectral index
model.  To understand where most measurements fall on this spectrum we
numerically quantify the overall model improvement derived from the addition of
the PAPER data as 
\begin{align}
\textrm{Improvement} &= \left(\textrm{Precision increase}\right) \left(\textrm{Catalog agreement}\right)\\
&= \left(\frac{1}{Area(\textrm{PAPER})} - \frac{1}{Area(\textrm{Cat})}\right) \left( Area(\PAPER \cap \textrm{Cat}) \right)
\end{align}

@@

However, the majority of source 78\% have positive improvement index,
indicating strong confirmation of the extrapolated spectrum (see Figures
\ref{fig:SI_contour_1} - \ref{fig:SI_contour_4}).  Pictor A is in the middle of
this group  with an improvement of 0.39 ---only 14 sources show stronger
confirmation.  The flux model is $S150=$388 $\pm$ 9.4Jy, $\alpha=-0.77\pm0.02$.
The combination of several independent PAPER data points near 150MHz with
previous catalog measurements results in this estimated uncertainty of 2.4\%.
This uncertainty is dominated by agreement between the 2 Wills measurements
near 600MHz which are accurate to 2\% and the five PAPER measurements, which
are accurate to ~5\%.  This error estimate includes an estimate of accuracy of the overall flux scale,
as well as the source-source variation due to primary beam errors.

% XXX comparison with previous in-band results (danny 2011, williams 2012)
% work TODO

@@

Here we have provided a measurement of Pictor A with enough precision to confirm
a highly linear spectral index between 150 and 600MHz. We then used the same
filtered beam form method to measure the spectra of bright sources with similar
primary beam response. This primary beam limit is based on our measurements 
of the primary beam along slices of constant declination.

The measurements provided
here are the first, calibrated, broad-band spectra to cover the EoR band. Existing EoR band measurements
are accurate to 20\% implying a 40\% uncertainty in the absolute power spectrum
level.  The Pictor A spectrum here is shown to be accurate to 
$\sim$5\%, a factor of 5 improvement over previous measurements.

This uncertainty includes the 
variation in each PAPER measurement ($\sim$1\%), variation between sources and the errors resulting
from extrapolating the \citet{Baars:1977p9678} scale beyond its original range. These last
two are found simultaneously by fitting the spectral extrapolations of several flux calibrators
at once and account for the the majority of the error. Though past measurements suggested the possibility
of spectral curvature below 200MHz we have found no evidence for this.  With these measurements we are able 
to confirm single spectral index model for Pictor between 120MHz and 600MHz.

A set of  61 verification sources were selected to be bright, previously cataloged
and  distributed over
the full 24 hours of Right Ascension coverage but only within 5 degrees
declination of Pictor A.  Using a Bayesian analysis we conclude that most of
these are consistent with previous measurements and provide useful new
constraints. However, in a small remainder, the simple spectral model fails to
adequately describe the data. In most cases the proximity to bright or extended
structures implies side lobe contamination.  Therefore, though we include the
model parameters, they are not meant to be definitive, but provide confidence
that the reported PAPER flux values are on the correct global scale.

Direct measurements of the Pictor A spectrum are key to correctly setting the
flux scale of PAPER, MWA and future EoR experiments like the Hydrogen Epoch of Reionization Array (HERA).
These spectra provide tighter constraints on many of the EoR band fluxes, while
limiting the pernicious affect of primary beam uncertainty.  Future work will use these
fluxes to further  refine the primary beam models of these experiments which is crucial to properly reconstructing
both image and power spectrum flux.
